---
title: "Exercise based on data set cushny"
author: Xi Jin
format: html
editor: visual
---

In this exercise, I will use data set called "cushny" in package "psychTools". It is a small data set, so the results might have bias.

## Data import

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r, warning=FALSE}
#install.packages("psychTools")
library(psychTools)
df <- psychTools::cushny
head(df)

```

## Data description

```{r}
str(df)
summary(df)
sum(is.na(df)) #no missing data
#mean
mean_ct <-mean(df$Control) 
mean_d1 <- mean(df$drug1)
#median of control group,drug1
median_ct <- median(df$Control)
median_d1 <- median(df$drug1)
#quartiles
quart_ct <- quantile(df$Control, probs = c(0.25, 0.5, 0.75))
quart_d1 <- quantile(df$drug1, probs = c(0.25, 0.5, 0.75))
#sample variance and sample standard deviation
var(df$Control)
sd(df$Control)
var(df$drug1)
sd(df$drug1)
#range
range(df$Control)
range(df$drug1)
#IQR
IQR(df$Control)
summary(df$Control)

```

## Plots

```{r}
# In base R
hist(df$Control, breaks =3,freq = FALSE,
    main = "Distribution of control group")
# We can add mean (red) and median (blue): abline draws a straight line from a to b 
# (or vertical as here)
abline(v = mean_ct, col = "red") 
abline(v = median_ct, col = "blue")

#box plot
myseq <- as.character(c("control","drug1","drug2L", "drug2R","delta1", "delta2L","delta2R"))
boxplot(list(df$Control,df$drug1,df$drug2L,df$drug2R,df$delta1,df$delta2L,df$delta2R), ylab = "Hours of sleep",
        xlab = "group", names=myseq)
```

## T-distribution (exact CI)

```{r}

#SE of different groups, we know that the length of each group is 10,so df=9
n <- length(df$Control)
SE_ct <- sd(df$Control) / sqrt(10) 
SE_d1 <- sd(df$drug1) / sqrt(10) 



#t-distribution (exact CI)
str(df)

tQuantile <- qt(p = 0.975, df =  9)

#control group
lowerCI_mean_ct <- mean_ct - tQuantile * SE_ct

upperCI_mean_ct <- mean_ct + tQuantile * SE_ct

cbind(mean_ct, lowerCI_mean_ct, upperCI_mean_ct)

t.test(df$Control, data = df, var.equal = FALSE)


#drug1 group
lowerCI_mean_d1 <- mean_d1 - tQuantile * SE_d1

upperCI_mean_d1 <- mean_d1 + tQuantile * SE_d1

cbind(mean_d1, lowerCI_mean_d1, upperCI_mean_d1)
```

## CIs for difference in group means

```{r}
diff_mean_d1c <- mean_d1-mean_ct
var_diff_mean_d1c <- 
  (9 * var(df$drug1) + 9 * var(df$Control)) / 
  (9 + 9 - 2)
SE_diff_mean_d1c <- sqrt(var_diff_mean_d1c) * sqrt(1/9 + 1/9) 

tQuantile_diff <- qt(p = 0.975, df = 9 + 9 - 2)

lowerCI_diff_mean_d1c <- diff_mean_d1c - tQuantile_diff * SE_diff_mean_d1c
upperCI_diff_mean_d1c <- diff_mean_d1c + tQuantile_diff * SE_diff_mean_d1c
cbind(diff_mean_d1c, lowerCI_diff_mean_d1c, upperCI_diff_mean_d1c)
#The mean drug1 group sleep 0.75 hour than control group. 
#We are 95% confident that the true difference in the drug group lies between-1.31 hour and 2.81 hour.


```

## A paired t-test and a two group t-test

```{r}
#H0:the difference in the control group and drug1 group is zero
t.test(df$Control, df$drug1, data = df, var.equal = FALSE)
t.test(df$Control, df$drug1, data = df, var.equal = TRUE)
#p=0.4007>0.05, acceptH0, which shows there is no significant difference between the control group and drug1 group.
t.test(df$Control,df$drug2L, data = df, var.equal = FALSE)
t.test(df$Control,df$drug2L, data = df, var.equal = TRUE)
#p=0.0072647<0.05, rejuectH0, which shows there is significant difference between the control group and drug2L group.
t.test(df$Control,df$drug2R, data = df, var.equal = FALSE)
t.test(df$Control,df$drug2R, data = df, var.equal = TRUE)
#p=0.01155<0.05, rejuectH0, which shows there is significant difference between the control group and drug2L group.

```

## 
